---
title: "Wikipedia references"
author: "Wenceslao Arroyo-Machado"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyr)
library(dplyr)
```

## Preview

There are 29,276,667 Wikipedia mentions, of which 27,212,084 mentions have some identifier.

```{r import}
df <- read.csv(file='references/ref_data.csv', stringsAsFactors=FALSE)
df2 <- read.csv(file='references/url_ref_norm.tsv', stringsAsFactors = FALSE, sep='\t')
length(which(df$URL != '' | df$ID_list != ''))
```

Removing web archives.
```{r}
df$URL <- gsub('http[s]{0,1}://(web\\.archive|waybackmachine)\\.org/.*http', 'http', df$URL, perl=TRUE)
df$URL <- gsub('http[s]{0,1}://(web\\.archive|waybackmachine)\\.org/.*www', 'www', df$URL, perl=TRUE)
df$URL <- gsub('http[s]{0,1}://(web\\.archive|waybackmachine)\\.org/web/([0-9a-z]*|[0-9a-z]*\\*|\\*)/|http[s]{0,1}://(web\\.archive|waybackmachine)\\.org/\\*/', '', df$URL, perl=TRUE)
df$URL <- gsub('http[s]{0,1}://archive\\.(today|is|fo|li|vn|md)/.*http', 'http', df$URL, perl=TRUE)
df$URL <- gsub('http[s]{0,1}://archive\\.(today|is|fo|li|vn|md)/.*www', 'www', df$URL, perl=TRUE)
df$URL <- gsub('http[s]{0,1}://((www\\.)?)webcitation\\.org/.*http', 'http', df$URL, perl=TRUE)
df$URL <- gsub('http[s]{0,1}://((www\\.)?)webcitation\\.org/.*www', 'www', df$URL, perl=TRUE)
```

```{r}
all(df2$URL %in% df$URL)
sum(df$URL %in% df2$URL)
```

Fix URLs.
```{r}
url_n <- setNames(df2$URL_n, df2$URL)
url_n['https://archive.org/stream/appletonsillustr00newy#page/100/mode/2up\\n']
df$URL <- url_n[df$URL]
sum(!is.na(df$URL))
rm(df2)
```


## Identifiers

The ID_list field include all identifiers included in the reference.

```{r identifiers_preview}
head(df$ID_list[which(df$ID_list != '')], n = 10)
```

There are 20 kind of identifiers. ISBN and DOI are the most frequent. But the ISSN is not useful for the heterogeneous coupling at this paper-level. 

MR is MathSciNet review, ASIN is Amazon Standard Identification Number, OL is Open Library, ZBL is zbMATH, OSTI is Office of Scientific and Technical Information, JFM is Zentralblatt MATH (previous to ZBL), RFC is Request for Comments.

```{r identifiers}
df$ID_list <- gsub(pattern='^\\{|\\}$', replacement='', x=df$ID_list)

df_id <- tidyr::separate_rows(data=df[,c('id', 'ID_list')], ID_list, sep=', ')
df_id <- df_id[which(df_id$ID_list != ''),]

ids <- as.data.frame(table(gsub(pattern='=.*', replacement='', x=df_id$ID_list)), stringsAsFactors=FALSE)
ids <- ids[order(ids$Freq, decreasing = TRUE),]
head(ids, n=20)
```

## Publications and resources

```{r publications_dataframe}
df_pub <- unique(df[which(df$ID_list != ''), c('id', 'ID_list', 'URL')])
df_pub$pub_id <- 1:dim(df_pub)[1]

df_pub_ids <- df_pub[,c('id', 'pub_id', 'ID_list')]
names(df_pub_ids)[3] <- 'r_id'
df_pub_ids <- df_pub_ids[which(df_pub_ids$r_id != ''),]
df_pub_ids <- tidyr::separate_rows(df_pub_ids, r_id, sep = ', ')
df_pub_ids$type <- gsub(pattern = '\\=.*$', replacement = '', x=df_pub_ids$r_id)
df_pub_ids$r_id <- gsub(pattern = '^.*\\=', replacement = '', x=df_pub_ids$r_id)

# general trasnformations
# to lower
df_pub_ids$r_id <- tolower(df_pub_ids$r_id)
# remove spaces
df_pub_ids$r_id <- gsub(pattern = '\\s', replacement = '', x=df_pub_ids$r_id)
# remove escape characters
df_pub_ids$r_id <- gsub(pattern='(\\\\n)+$|(\\n)+$|(\\\\t)+$|(\\t)+$', replacement='', x=df_pub_ids$r_id)
df_pub_ids$r_id <- gsub(pattern='^(\\\\n)+|^(\\n)+|^(\\\\t)+|^(\\t)+', replacement='', x=df_pub_ids$r_id)


# ISBN transformations
df_pub_ids$r_id[which(df_pub_ids$type == 'ISBN')] <- gsub(pattern = '-', replacement = '', x=df_pub_ids$r_id[which(df_pub_ids$type == 'ISBN')])

table(nchar(df_pub_ids$r_id[which(df_pub_ids$type == 'ISBN')]))
df_pub_ids$r_id[which((!(nchar(df_pub_ids$r_id) %in% c(10, 13))) & df_pub_ids$type == 'ISBN')] <- ''

table(nchar(df_pub_ids$r_id[which(df_pub_ids$type == 'DOI')]))
table(substr(df_pub_ids$r_id[which(df_pub_ids$type == 'DOI')], 1, 4))
df_pub_ids$r_id[which((!(grepl(pattern = '^10', x = df_pub_ids$r_id))) & df_pub_ids$type == 'DOI')] <- ''
df_pub_ids <- df_pub_ids[which(df_pub_ids$type != 'ISSN'),]

df_pub_ids <- df_pub_ids[which(df_pub_ids$r_id != ''),]



df_pub_url <- df_pub[,c(1,4,3)]
names(df_pub_url)[3] <- 'r_id'
df_pub_url <- df_pub_url[which(!is.na(df_pub_url$r_id)),]
df_pub_url$type <- 'URL'

df_pub <- rbind.data.frame(df_pub_url, df_pub_ids, stringsAsFactors = FALSE)
```


```{r publications_dataframe}
df_pub <- unique(df[which(df$URL != '' | df$ID_list != ''), c('ID_list', 'URL')])
df_pub$pub_id <- 1:dim(df_pub)[1]

df_pub_ids <- df_pub[,c(3,1)]
names(df_pub_ids)[2] <- 'r_id'
df_pub_ids <- df_pub_ids[which(df_pub_ids$r_id != ''),]
df_pub_ids <- tidyr::separate_rows(df_pub_ids, r_id, sep = ', ')
df_pub_ids$type <- gsub(pattern = '\\=.*$', replacement = '', x=df_pub_ids$r_id)
df_pub_ids$r_id <- gsub(pattern = '^.*\\=', replacement = '', x=df_pub_ids$r_id)

# general trasnformations
df_pub_ids$r_id <- tolower(df_pub_ids$r_id)


df_pub_ids$r_id <- gsub(pattern = '\\s', replacement = '', x=df_pub_ids$r_id)
df_pub_ids$r_id <- gsub(pattern='(\\\\n)+$|(\\n)+$|(\\\\t)+$|(\\t)+$', replacement='', x=df_pub_ids$r_id)
df_pub_ids$r_id <- gsub(pattern='^(\\\\n)+|^(\\n)+|^(\\\\t)+|^(\\t)+', replacement='', x=df_pub_ids$r_id)

df_pub <- df_pub_ids
# review duplicated
df_pub <- unique(df_pub[which(df_pub$r_id != ''), c('pub_id', 'r_id', 'type')])



# ISBN
df_pub_ids$r_id[which(df_pub_ids$type == 'ISBN')] <- gsub(pattern = '-', replacement = '', x=df_pub_ids$r_id[which(df_pub_ids$type == 'ISBN')])

df_pub_ids$r_id[which((!(nchar(df_pub_ids$r_id) %in% c(10, 13))) & df_pub_ids$type == 'ISBN')] <- ''

table(nchar(df_pub_ids$r_id[which(df_pub_ids$type == 'DOI')]))
df_pub_ids$r_id[which((!(grepl(pattern = '^10', x = df_pub_ids$r_id))) & df_pub_ids$type == 'DOI')]


df_pub_url <- df_pub[,c(3,2)]
names(df_pub_url)[2] <- 'r_id'
df_pub_url <- df_pub_url[which(df_pub_url$r_id != ''),]
df_pub_url$type <- 'URL'

df_pub <- rbind.data.frame(df_pub_url, df_pub_ids, stringsAsFactors = FALSE)
df_pub <- df_pub[which(df_pub$type != 'ISSN'),]
# remove worngs ids
df_pub <- df_pub[which(df_pub$type %in% c(ids$Var1[1:20],'URL')),]

# rename page_id
dict <- setNames(as.character(1:length(unique(df_pub$pub_id))), as.character(unique(df_pub$pub_id)))
df_pub$pub_id <- dict[as.character(df_pub$pub_id)]
df_pub$pub_id <- as.integer(df_pub$pub_id)

write.table(x=df_pub, file='references/wp_ref.tsv', quote = FALSE, row.names = FALSE, sep = '\t', fileEncoding = 'UTF-8')
```
